\protect Connecting Every Bit of Knowledge: \\ Wikipedia's First Link Network Unravelled

### 2015-12-09 writeup/2015-11-wikipedia-network.tex
directory cleanup
fixed figure placement

\begin{abstract}
  \protect
Apples, oranges, and the most obscure Dylan song too---is everything a few clicks from Philosophy? 
Within Wikipedia, the surprising answer is yes: nearly all 
paths lead to Philosophy.
Wikipedia is the largest, most meticulously indexed collection of human knowledge ever amassed. 
More than information about a topic though, Wikipedia is a marvelous web of naturally emerging relationships.  
By following the First Link in an article, we connect entries to form a directed network within Wikipedia: Wikipedia's First Link Network. 
Here we study the English edition of Wikipedia's First Link Network for insight into how we relate topics, ideas, people, objects, and events.

We algorithmically parse all 4.7 million articles to construct a map of Wikipedia's First Link Network. 
In a novel approach to uncover structure, we traverse every possible path through the network, 
measuring the accumulation of First Links, path lengths, basins, cycles, and even particular articles funneling links into the cycles.
We discover many scale-free distributions, find Philosophy at a salient center, and uncover a flow from specific to general with 
basins around fundamental notions such as Community, State, and Science. 
Curiously, we also observe a gravitation towards topical articles including Health Care and Fossil Fuel.
These findings enrich our view of how we connect and structure
an ever growing load of information.

@@

It's not surprising then to see Wikipedia as the object of many studies. 
Researchers have examined the cultural dynamics among editors,
\cite{editors}
the accuracy of the content relative to traditional encyclopedias,
\cite{accuracy1}
\cite{accuracy2}
the disciplines and topics covered 
\cite{coverage}
and bias against portions of the population
\cite{bias_women}.
Wikipedia's content has also proven a powerful tool. 
Researchers have used Wikipedia to identify missing dictionary entries
\cite{missing_entries},
cluster short text
\cite{clustering},
compute semantic relatedness
\cite{semantic_relatedness},
and disambiguate meaning \cite{disambiguating}.

@@

The authors of a Wikipedia article choose where and whether to include a 
hyperlink reference to another Wikipedia article
by including hyperlinks in the article's markup.
In describing a particular topic, each set of authors decides which articles are the most
relevant references.
For example, the hyperlinks in the "Train" article depend only on what its authors 
decide are pertinent articles to reference.
The collection of hyperlinks emerges from the independent choice of each article's authors. 
The first hyperlink to another Wikipedia article marks the earliest moment in the body text where the authors
choose to connect the information discussed to that of another article.
By following the first hyperlink to another article in the English edition of
Wikipedia, we connect the ideas in one article to those in another, ultimately forming a directed network: 
{\it Wikipedia's First Link Network} (FLN).

\begin{figure*}[tp!]
  \centering	
  \includegraphics[width=\textwidth]{graphics/Train.pdf}  
  \caption{
    \textbf{First Link Path For "Train"}
    We follow the first link to another Wikipedia article
    in the main body of the article---the area
    inside the green rectangle, which excludes 
    side bar elements, the navigation bar and title. 
    In this example, the first link to another Wikipedia article is "Rail Transport." 
    We can again select the first link on the "Rail Transport"
    article, repeating the process to 
    form a path of First Links.
  }
  \label{fig:Train First Links}
\end{figure*}

What information could the first links possibly reveal?
Inspired by the claim that the majority of first links lead to the 
"Philosophy" article---popularized by an xkcd comic and subsequently
discussed in blog posts 
\cite{mat_blog}
\cite{Ilmari_first_links}
\cite{xkcd} 
---we holistically study 
Wikipedia's FLN as a map relating areas of human knowledge. 
The map is a hierarchical taxonomy where "Train" links to a parent node, 
"Rail Transport," while many child nodes, "Steel" and "Horsepower," link up to 
"Train." Unlike previous taxonomies
created by individuals
\cite{locke}
\cite{descartes}
\cite{aristotle} 
or a select group of individuals 
\cite{hist_thesaurus}, 
the organization of ideas in the FLN 
does not emerge from a concerted centralized effort, 
but rather from the independent choice of every article's authors.
The FLN is a wealth of relations among all the inventions, places,
figures, objects, and events described in Wikipedia, across space and time.

Our goal is to gain insight into how the information on Wikipedia is organized and related
by studying the structure of the FLN.
In a novel approach, we develop metrics to capture 
the dominant features of the FLN's structure.
We measure dynamics of the FLN as a flow, quantifying 
aspects of the FLN from the accumulation of first links around articles 
to the influence an article exerts in shaping the FLN.
Together with cycles, indegree, depth, and the content of the articles, 
we build our analysis of the relations among the ideas in Wikipedia.

@@

One essential measure of a directed network structure is the degree distribution 
\cite{newman}. 
Degree distribution has been used to study many phenomena from disease outbreak 
\cite{disease} 
to the dynamics of social networks 
\cite{social_nets}.
The degree distribution in the FLN describes how many first links point to a particular article. 
We measure the indegree in the FLN to understand which articles authors choose to reference in 
the first link.
A higher indegree indicates more authors selected the particular article
as a relevant reference. On the other hand, articles with zero 
indegree have no references---they are outer leaves in the FLN. 
Similar to PageRank, the indegree is also a way to rank the articles in Wikipedia.
We can rank articles by indegree to see which articles are most and least referenced in the FLN. 
Additionally, we can assess the degree distribution to see whether authors tend to reference all articles
equally or only a few, as observed in many social networks \cite{social_nets}.

The degree distribution captures one step in the FLN, measuring   
the destination of a given link. 
Indegree measures only the particular set of first links between two articles, 
rather than the richer dynamics of how many articles fit together---how links 
create a flow through the FLN.
One way to expand the analysis beyond a single step is to map the full path of 
an article's first link. 
Each article is both a destination and a link to another article, similar to an element 
in a linked list data structure. 
By following the chain of links we can construct a path through the FLN with each article 
leading to the next via the first link. 
For example "Train" has a first link to "Conveyance of Passengers and Goods," which is itself
an article with a first link to "Goods" and so on. The path starting at "Train" 
is the collection of articles obtained by following the first links starting at "Train"
(see figure~\ref{fig:Train First Links}).
The path contains "Goods," "Economics," "Social Science," and other articles related to 
"Train," but not directly referenced by a first link.

Applying a similar concept with a 
different starting article, we form 
a path through the FLN moving from one article to the next.
A path in the FLN connects a group of articles allowing us to measure how two articles relate,
even if the articles are not directly linked. 
The collection of articles is path-connected conveying a flow of concepts from one article to another. 
The notion of paths allows us to map not only which articles link directly to "Philosophy" 
for example, but also which ideas are eventually anchored in "Philosophy" even if the articles
are not directly linked.
The method is order agnostic with respect to which articles are selected first. As long as each article is selected eventually, the resulting metrics are equivalent.
Starting at each article, we construct a path through the FLN for each article, 
ultimately mapping how the first links flow through the FLN.

In a novel approach, we use the notion of a path through the FLN to describe 
the structure of the FLN as a flow relating ideas. 
Grounded in the notion of flow through river networks,
\cite{geo_basins}
we characterize features of the FLN as a flow via first links.
Previous studies have used flow to characterize the structure of river networks
\cite{dodds} and describe the organization of food systems as transportation networks
\cite{food_webs}.
With paths to mark flow through the FLN, we develop new metrics to 
gauge the accumulation of references, 
the length of the path relating articles, and the influence a particular
article exerts in shaping the flow of links through the FLN.

@@

Our goal is to map the flow of first links by following the paths of first links 
starting at each article. 
Since a path reveals how the ideas in one article may eventually be 
anchored in another, the first metric we develop quantifies the accumulation of first links 
within the FLN. 
The algorithm begins by selecting an article then traversing the path formed
by following the first links. Each time a first link references an article, we increment a count
associated with the article. 
We repeat until the first link is repeated or invalid (outside the FLN). 
The collection of articles is path-connected conveying a flow of concepts from one article to another. 
We select a second article and repeat the process until we have 
constructed a path for each article in the network. We call the resulting count associated with each article the number of {\it traversal visits}. The number of traversal visits of an article 
measures the number of references flowing to the ideas in the article.

We can map the paths in the FLN as a matrix with each column corresponding to a path. In our sample network 
(see figure~\ref{fig:Traversal Visits}), the path starting at article A is the 
first column in the traversal vists matrix with $1$ if the path contains a given article and 
$0$ if the path doesn't. The second column indicates the path starting at article B and so on. 
To compute the number traversal visits for article A we sum the corresponding row in our matrix,
which totals the number of paths containing article A. The total is analogous to the flow 
in a river network with traversal visists measuring the accumulation of water near point A. 
By constructing a traversal visits matrix, we can similarly compute the number of traversal 
visits by summing along the corresponding row for our node. The traversal visits matrix for 
Wikipedia consists 121 million entries to account for each path through the FLN. 
We measure the accumulation of first link references by summing the corresponding row for the 
article of interest.

In addition to accumulation, the traversal visits algorithm creates paths 
which uncover depth in the FLN. 
By measuring the number of first links between
two articles, we get an additional piece of information we call 
{\it path length.} We can compute the length a path by summing 
along a column in our traversal visits matrix. 
The path length describes how closely related ideas are, gauging the FLN's depth.
The greater the number of first links separating two articles, the greater the number of ideas needed to relate the ideas. Although, 
"Train" is related to "Economics," there are several more specific articles
bridging the connection:
"Train" is more specifically related to transportation, whose object
is often goods. Goods are one of the fundamental objects of study in Economics.
Described in links, this relationship is captured by a relatedness of $4$ 
first links ultimately connecting "Train" to "Economics." 
We can examine path length as a relative measure of how closely connected ideas
may be, but also as an aggregate measure describing the overall separation or 
connectedness in the FLN.
We can characterize whether ideas a closely connected by a handful of links, 
many, or whether a more nuanced organization exists.

One possible path through the FLN is a {\it cycle} or a group of articles 
linking to one another inside a loop. In our sample network 
(see figure~\ref{fig:Traversal Visits}) 
a cycle exists among nodes A, B, and C. 
Since node C links back to the starting node A, the path starting at node A forms a {\it 3-cycle}. 
By recording the history of the articles along a path in the FLN, we can identify
the types of cycle structures within the FLN. Furthermore, since each article 
in the a cycle has an associated traversal visit count, we can rank cycles by 
the number of references directed towards a cycle. 
We can also form and rank {\it basins} in the network by identifying groups of 
path-connected articles, not necessarily forming a perfect cycle.
A basin connects a group of articles which ultimately reference other articles
in the FLN. By identifying basins and cycles, we pinpoint group of closely related articles.

@@

While traversal visits measure accumulation, each article's first link also 
influences the shape of the FLN. 
At a large point of accumulation, a single article's first link 
can exert great influence over the shape of the FLN by directing many
references towards a particular path. To distinguish between an article 
that simply happened to fall within a cycle from an article funneling 
many first links, we develop a second metric called {\it traversal funnels}.

To measure traversal funnels, we traverse the FLN in the same manner as we 
did for traversal visits, but end a path once we enter a cycle.
We are then able to distinguish between an article related to many other ideas
only by virtue of its place in a cycle from an article exerting influence over where the first links flow. 
An article with a large number of traversal funnels directs many references
towards a particular path. In our sample network 
(see figure~\ref{fig:Traversal Funnels}) article C 
directs the flow of links towards the 3-cycle, while articles A and B are 
receipients of the flow---without exerting any influence themselves. 
While articles A, B, and C all have the same number of traversal visits, 
article C has 4 traversal funnels (A and B have none). By incrementing the count of 
first link references on paths only up to a cycle, we can identifying which articles
exert the greatest influence over the FLN's structure.

@@

We first identify the set of first links directly referencing a particular article 
by studying the indegree of the FLN. 
A higher indegree suggests more authors found the ideas in the article 
a pertinent direct reference. 
\begin{figure*}[tp!]
  \centering	
  \includegraphics[width=\columnwidth]{graphics/articles_ndegree.png}
  \caption{
    \textbf{Highest Ranking Articles by indegree}
  }
  We rank each article by the number of direct first links to the article (indegree). The highest-ranking articles tend to represent abstraction for concepts
  comprised of various facets and parts.
  \label{fig:ndegree list}
\end{figure*}
We rank all 11 million articles by indegree to find 
the "United States" with 80,249 direct first links as the most referenced
Wikipedia article. 
Other high-ranking articles
include foundational abstract concepts such as "village," "species,"; 
sports associations such as "American Football," "Association Football"; 
and developed nations such as "France," "Japan," "Russia," "Australia," and 
the "Netherlands." These high ranking articles are useful abstractions: nations
describe a collection of individuals, a common culture, language, or 
geographical proximity; sports teams describe an ever changing collection of 
sports players often associated with a cultural identity and a geographical 
region. Since abstractions such as nations and teams are inherently comprised
of many parts, authors reference the larger abstraction when describing 
a part. For example, in describing the geographical region of north America, 
the United States, Canada, or Mexico are natural references anchoring the 
geographical location to a larger national identity.

"Philosophy" and other philosophical concepts with many traversal visits
are not among the highest-ranking articles by indegree.
"Philosophy" has an indegree of 581, with direct first links from articles about Philosophers and areas of Philosophy: "Existentialism and Humanism," "Predeterminism," "Synoptic Philosophy,"Qualia," "Dorothy Emmet," and "Christopher W. Morris."
While many articles accumulate at "Philosophy" (see traversal visits discussion), 
the accumulation is not the 
result of many articles directly referencing "Philosophy." 
Instead, the accumulation of first links, as we argue in our 
discussion of traversal visits, flows towards Philosophy as the 
ultimate anchor when genralizing from specific articles towards broader notions.
\begin{figure*}[tp!]
  \centering	
  \includegraphics[width=\columnwidth]{graphics/ndegree_loglog.png}
  \caption{
    \textbf{FLN Degree Distribution}
  }
  We construct a log-log plot fit our results to a linear model. The result is a 
  an excellent fit with r = -0.98, yielding a power law exponent of -0.79. 
  The distribution appears to be scale free with most articles having fewer than 9 
  direct first links, while few hold most.
  \label{fig:degree distribution}
\end{figure*}
The FLN's indegree exhibits a scale-free distribution where a few articles 
receive a most direct first references, while most articles receive few or none.
The average indegree for all 11 million articles is 3.6 direct first links with a standard deviation of 89.5 links.
Only 4,826 articles have more than 100 direct first links and $75\%$ of articles
have fewer than 9. 
When fit to a linear model on a log-log scale (log(rank) versus log(indegree)), 
the model's r value is $-0.98$ suggesting a strong log-log linear fit 
with a power law exponent of $-0.79$.

@@

\begin{figure}[tp!]
  \centering	
  \includegraphics[width=\columnwidth]{graphics/path_lengths_boxplot.png}
  \caption{
    \textbf{Path Length Distribution}
  }
  The median network depth measured by the number of first links in a path
  is 29. The 365 cycle for "Orthodox" liturgics is the outlier to the right, while
  other historical articles about "Scotland," "UK" and so on are slightly outside
  the third quartile. More than $75\%$ of articles have path lengths between 
  $0$ and $50$ links.

@@

Of the 11 million articles, 5.5 million had an invalid link or linked back to the same article, yielding a path length of zero. 
This roughly corresponds to the official number of articles on Wikipedia 
$~4.7$ million as of November 2014---approximately half of the 11 million 
articles in the XML dump are redirects or disambiguations, not full articles.
The most common path length is 29, with an interquartile range (26, 30).
The median path length far exceeds the popular 6-degrees of separation (see for 
example the average number of academic publications separating scholars 
\cite{six_degrees}), indicating a greater network depth. The network depth suggests a large network of spanning a variety of ideas.
As a distribution, more than $75\%$ of articles have a path length below 
$50$ first links 
while a few temporally organized paths exceed 50 links 
(see figure~\ref{fig:Path Length Distribution}).

How can we characterize the flow of ideas along a path? Next, we develop 
a metric to gauge the accumulation first link references
and characterize a global organization 
of ideas among articles.

@@

We followed every possible path through the network, incrementing
the number of traversal visits for every first link reference along a path 
to generate a total of 232 million traversal visits.

As a distribution, the number traversal visits per article appears to be scale-free. The majority of articles have fewer than 30 traversal visits, while few 
have 5 order of magnitude more traversal visits. 
Specifically, $99.76\%$ of articles have fewer than $100$ traversal visits; nearly $80\%$ have none. 
Meanwhile, the highest ranking 30 articles have an extremely disproportionate number of traversal visits.

\begin{figure*}[tp!]
  \centering	
  \includegraphics[width=\columnwidth]{graphics/traversals_per_article.png} 
  \caption{
    \textbf{Distribution of Traversal Visits}
  }
  We fit the distribution to a linear log-log model by considering the log (base 10) transformed rank of each article against log (base 10) transformed the number of traversal visits. 
  The model explains $86\%$ of the variation in the data, yielding an r-value of $-0.93$ 
  and a power law exponent of $-0.636$. The horizontal flattening around the highest
  ranking articles is a result of the cyclic structure (see discussion on cycles).
  \label{fig:Distribution of Visits}

@@

The highest ranking articles include Philosophy alongside related articles such as "Existence", "Quality", and "Reality".

\begin{figure*}[tp!]
  \centering	
  \includegraphics[width=\columnwidth]{graphics/articles_ranked.png}
  \caption{
    \textbf{highest ranking articles by number of traversal visits}
  }
  We compute the number of traversal visits for each article in the FLN (see 
  Traversal Algorithm section for details). In doing so, we can rank each article
  by the accumulation of first links. Articles with a greater number of traversal visits
  mark greater points of first link accumulation. The highest ranking articles by traversal visits reveals where the greatest accumulation occurs.
  \label{fig:highest visits}
\end{figure*}

@@

\begin{figure*}[tp!]
  \centering	
  \includegraphics[width=\columnwidth]{graphics/synonyms.png}
  \caption{
    \textbf{Number of Synonyms for the highest-ranking articles}
    We use WordNet (the largest lexical database for the English language)
    to identify the number of synonyms for an article's title. We compare 
    the distribution of synonyms of the highest-ranking 100 articles
    against all other Wikipedia articles. As the distribution demonstrates
    the highest ranking concepts have more synonyms than other articles.
  }
  \label{fig:synonyms}
\end{figure*}

>>>>

\begin{abstract}
  \protect
Apples, oranges, and the most obscure Dylan song too---is everything a few clicks from Philosophy? 
Within Wikipedia, the surprising answer is yes: nearly all 
paths lead to Philosophy.
Wikipedia is the largest, most meticulously indexed collection of human knowledge ever amassed. 
More than information about a topic though, Wikipedia is a web of naturally emerging relationships.  
By following the First Link in an article, we connect entries to form a directed network within Wikipedia: Wikipedia's First Link Network. 
Here we study the English edition of Wikipedia's First Link Network for insight into how the many topics, ideas, people, objects, and events
are related and organized.

We algorithmically parse all 4.7 million articles to construct a map of Wikipedia's First Link Network. 
In a novel approach to uncover structure, we traverse every possible path through the network, 
measuring the accumulation of First Links, path lengths, basins, cycles, and even the influence a particular article exerts in shaping the 
network.
We discover many scale-free distributions, find Philosophy at a salient center, and uncover a flow from specific to general with 
culminating around fundamental notions such as Community, State, and Science. 
Curiously, we also observe a gravitation towards topical articles including Health Care and Fossil Fuel.
These findings enrich our view of how we connect and structure
an ever growing load of information.

@@

It's not surprising then to see Wikipedia as the object of many studies. 
Researchers have examined the cultural dynamics among editors
\cite{editors},
the accuracy of the content relative to traditional encyclopedias
\cite{accuracy1},
\cite{accuracy2},
the disciplines and topics covered 
\cite{coverage},
and bias against portions of the population
\cite{bias_women}.
Wikipedia's content has also proven a powerful tool. 
Researchers have used Wikipedia to identify missing dictionary entries
\cite{missing_entries},
cluster short text
\cite{clustering},
compute semantic relatedness
\cite{semantic_relatedness},
and disambiguate meaning \cite{disambiguating}.

@@

The authors of a Wikipedia article choose where and whether to include a 
hyperlink reference to another Wikipedia article
by including hyperlinks in the article's markup.
In describing a particular topic, the authors decide which other articles are most
relevant references.
For example, the authors of the "Train" article chose
"Rail Transport," "Cargo," and "Passengers" 
as pertinent articles to reference in describing trains.
The collection of hyperlinks emerges from the independent choice of each article's authors. 
The first hyperlink to another Wikipedia article marks the earliest moment in the body text where the authors
choose to connect the information discussed to that of another article.
By following the first hyperlink to another article in the English edition of
Wikipedia, we connect the ideas in one article to those in another, ultimately forming a directed network: 
{\it Wikipedia's First Link Network} (FLN).

\begin{figure*}[tp!]
  \centering	
  \includegraphics[width=\textwidth]{graphics/Train.pdf}  
  \caption{
    \textbf{First Link Path For "Train"}
    We follow the First Link to another Wikipedia article
    in the main body of the article---the area
    inside the green rectangle, which excludes 
    side bar elements, the navigation bar and title. 
    In this example, the First Link to another Wikipedia article is "Rail Transport." 
    We can again select the First Link on the "Rail Transport"
    article, repeating the process to 
    form a path of First Links.
  }
  \label{fig:Train First Links}
\end{figure*}

What information could the First Links possibly reveal?
Inspired by the claim that the majority of First Links lead to the 
"Philosophy" article---popularized by an xkcd comic and subsequently
discussed in blog posts 
\cite{mat_blog}
\cite{Ilmari_first_links}
\cite{xkcd} 
---we holistically study 
Wikipedia's FLN as a map relating areas of human knowledge. 
The map is a hierarchical taxonomy where "Train" links to a parent node, 
"Rail Transport," while many child nodes, "Steel" and "Horsepower," link up to 
"Train." Unlike previous taxonomies
created by individuals
\cite{locke}
\cite{descartes}
\cite{aristotle} 
or a select group of individuals 
\cite{hist_thesaurus}, 
the organization of ideas in the FLN 
does not emerge from a concerted or a centralized effort. 
Rather each link in the FLN emerges from the independent choice of an article's authors.
Collectively, the FLN is a wealth of relations among inventions, places,
figures, objects, and events across space and time.

Our goal is to gain insight into how the information on Wikipedia is organized and related
by studying the structure of the FLN.
In a novel approach, we develop metrics to capture 
the dominant features of the FLN's structure.
We measure dynamics of the FLN as a flow, quantifying 
aspects of the FLN from the accumulation of First Links around articles 
to the influence an article exerts in shaping the FLN.
Together with cycles, indegree, depth, and the content of the articles, 
we build our analysis of the relations among the ideas in Wikipedia.

@@

One essential measure of a directed network structure is the degree distribution 
\cite{newman}. 
Degree distribution has been used to study many phenomena from disease outbreak 
\cite{disease} 
to the dynamics of social networks 
\cite{social_nets}.
The degree distribution in the FLN describes how many First Links point to a particular article. 
We measure the indegree in the FLN to guage the number of articles referencing a particular article.
A higher indegree indicates more authors selected the particular article
as a relevant reference. On the other hand, articles with zero 
indegree have no references---they are outer leaves in the FLN. 
Similar to PageRank, the indegree is also a way to rank the articles in Wikipedia.
We can rank articles by indegree to see which articles are most and least referenced in the FLN. 
Additionally, we can assess the degree distribution to see whether authors tend to reference all articles
equally or only a few, as observed in many social networks \cite{social_nets}.

The degree distribution captures one step in the FLN, measuring   
the destination of a given link. 
Indegree measures only the particular set of First Links between two articles, 
rather than the richer dynamics of how many articles fit together---how links 
create a flow through the FLN.
One way to expand the analysis beyond a single step is to map the full path of 
an article's First Link. 
Each article is both a destination and a link to another article, similar to an element 
in a linked list data structure. 
By following the chain of links we can construct a path through the FLN with each article 
leading to the next via the First Link. 
For example "Train" has a First Link to "Conveyance of Passengers and Goods," which is itself
an article with a First Link to "Goods" and so on. The path starting at "Train" 
is the collection of articles obtained by following the First Links starting at "Train"
(see figure~\ref{fig:Train First Links}).
The path contains "Goods," "Economics," "Social Science," and other articles related to 
"Train," but not directly referenced by a First Link.

@@

Applying a similar concept with a 
different starting article, we form 
a path through the FLN moving from one article to the next.
A path in the FLN connects a group of articles allowing us to measure how two articles relate,
even if the articles are not directly linked. 
The collection of articles is path-connected conveying a flow of concepts from one article to another. 
The notion of paths allows us to map not only which articles link directly to "Philosophy" 
for example, but also which ideas are eventually anchored in "Philosophy" even if the articles
are not directly linked.
The method is order agnostic with respect to which articles are selected first. As long as each article is selected eventually, the resulting metrics are equivalent.
Starting at each article, we construct a path through the FLN for each article, 
ultimately mapping how the First Links flow through the FLN.

In a novel approach, we use the notion of a path through the FLN to describe 
the structure of the FLN as a flow relating ideas. 
Grounded in the dynamics of river networks,
\cite{geo_basins}
we characterize features of the FLN as a flow through First Links.
Previous studies have used flow to characterize the structure of river networks
\cite{dodds} and describe the organization of food systems as transportation networks
\cite{food_webs}.
With paths to mark flow through the FLN, we develop new metrics to 
gauge the accumulation of references, 
the length of the path relating articles, and the influence a particular
article exerts in shaping the flow of links through the FLN.

Our goal is to map the flow of First Links by following the paths of First Links 
starting at each article. 
Since a path reveals how the ideas in one article may eventually be 
anchored in another, the first metric we develop quantifies the accumulation of First Links 
within the FLN. 
The algorithm begins by selecting an article then traversing the path formed
by following the First Links. Each time a First Link references an article, we increment a count
associated with the article. 
We repeat until the First Link is repeated or invalid (outside the FLN). 
We select a second article and repeat the process until we have 
constructed a path for each article in the network. We call the resulting count for each article the number of {\it traversal visits}. The number of traversal visits of an article 
measures the number of references flowing to the ideas in the article.

@@

We can map the paths in the FLN as a matrix with each column corresponding to a path. In our sample network 
(see figure~\ref{fig:Traversal Visits}), the path starting at article A is the 
first column in the traversal vists matrix. 
An entry of $1$ indicates the path contains a given article and 
$0$ indicates the path doesn't. The columns identify the starting article of the path. 
For example, the second column corresponds to the path starting at article B. 
To compute the number traversal visits for article A we sum the corresponding row in our matrix.
The total is the number of paths containing article A. 
Analogous to the flow 
in a river network, traversal visits measure the accumulation of water near point A. 
By constructing a traversal visits matrix, we can similarly compute the number of traversal 
visits by summing along the corresponding row for our node. The traversal visits matrix for 
Wikipedia consists 121 million entries to account for each path through the FLN. 
We measure the accumulation of First Link references by summing the corresponding row for the 
article of interest.

In addition to accumulation, the traversal visits algorithm creates paths 
which uncover depth in the FLN. 
By measuring the number of First Links between
two articles, we get an additional piece of information we call 
{\it path length.} We can compute the length a path by summing 
along a column in our traversal visits matrix. 
The path length describes how closely related ideas are, gauging the FLN's depth.
The greater the number of First Links separating two articles, the greater the number of ideas needed to relate the ideas. Although, 
"Train" is related to "Economics," there are several more specific articles
bridging the connection:
"Train" is more specifically related to transportation, whose object
is often goods. Goods are one of the fundamental objects of study in Economics.
Described in links, this relationship is captured by a relatedness of $4$ 
First Links ultimately connecting "Train" to "Economics." 
We can examine path length as a relative measure of how closely connected ideas
may be, but also as an aggregate measure describing the overall separation or 
connectedness in the FLN.
We can characterize whether ideas a closely connected by a handful of links, 
many, or whether a more nuanced organization exists.

One possible path through the FLN is a {\it cycle} or a group of articles 
linking to one another inside a loop. In our sample network 
(see figure~\ref{fig:Traversal Visits}) 
a cycle exists among nodes A, B, and C. 
Since node C links back to the starting node A, the path starting at node A forms a {\it 3-cycle}. 
By recording the history of the articles along a path in the FLN, we can identify
the types of cycle structures within the FLN. Furthermore, since each article 
in the a cycle has an associated traversal visit count, we can rank cycles by 
the number of references directed towards a cycle. 
We can also form and rank {\it basins} in the network by identifying groups of 
path-connected articles, not necessarily forming a perfect cycle.
A basin connects a group of articles which ultimately reference other articles
in the FLN. By identifying basins and cycles, we pinpoint groups of closely related articles.

While traversal visits measure accumulation, each article's First Link also 
influences the shape of the FLN. 
At a large point of accumulation, a single article's First Link 
can exert great influence over the shape of the FLN by directing many
references towards a particular path. To distinguish between an article 
that simply happened to fall within a cycle from an article funneling 
many First Links, we develop a second metric called {\it traversal funnels}.

To measure traversal funnels, we traverse the FLN in the same manner as we 
did for traversal visits, but end a path once we enter a cycle.
We are then able to distinguish between an article related to many other ideas
only by virtue of its place in a cycle from an article exerting influence over where the First Links flow. 
An article with a large number of traversal funnels directs many references
towards a particular path. In our sample network 
(see figure~\ref{fig:Traversal Funnels}) article C 
directs the flow of links towards the 3-cycle, while articles A and B are 
receipients of the flow---without exerting any influence themselves. 
While articles A, B, and C all have the same number of traversal visits, 
article C has 4 traversal funnels (A and B have none). By incrementing the count of 
First Link references on paths only up to a cycle, we can identifying which articles
exert the greatest influence over the FLN's structure.

@@

\begin{figure}[tp!]
  \centering	
  \includegraphics[width=\columnwidth]{graphics/articles_ndegree.png}
  \caption{
    \textbf{Highest Ranking Articles by indegree}
  }
  We rank each article by the number of direct First Links to the article (indegree). The highest-ranking articles tend to represent abstraction for concepts
  comprised of various facets and parts.
  \label{fig:ndegree list}
\end{figure}

We first identify the set of First Links directly referencing a particular article 
by studying the indegree of the FLN. 
A higher indegree suggests more authors found the ideas in the article 
pertinent to reference. 
We rank all 11 million articles by indegree to find 
the "United States" with 80,249 direct First Links as the most referenced
Wikipedia article. 
Other high-ranking articles
include foundational abstract concepts such as "village," "species,"; 
sports associations such as "American Football," "Association Football"; 
and developed nations such as "France," "Japan," "Russia," "Australia," and 
the "Netherlands." These high ranking articles are useful abstractions: nations
describe a collection of individuals, a common culture, language, or 
geographical proximity; sports teams describe an ever changing collection of 
sports players often associated with a cultural identity and a geographical 
region. Since abstractions such as nations and teams are inherently comprised
of many parts, authors reference the larger abstraction when describing 
a part. For example, in describing the geographical region of north America, 
the United States, Canada, or Mexico are natural references anchoring the 
geographical location to a larger national identity.

"Philosophy" and other philosophical concepts with many traversal visits
are not among the highest-ranking articles by indegree.
"Philosophy" has an indegree of 581, with direct First Links from articles about Philosophers and areas of Philosophy: "Existentialism and Humanism," "Predeterminism," "Synoptic Philosophy,"Qualia," "Dorothy Emmet," and "Christopher W. Morris."
While many articles accumulate at "Philosophy" (see traversal visits discussion), 
the accumulation is not the 
result of many articles directly referencing "Philosophy." 
Instead, the accumulation of First Links, as we argue in our 
discussion of traversal visits, flows towards Philosophy as the 
ultimate anchor when genralizing from specific articles towards broader notions.

\begin{figure}[tp!]
  \centering	
  \includegraphics[width=\columnwidth]{graphics/ndegree_loglog.png}
  \caption{
    \textbf{FLN Degree Distribution}
  }
  We construct a log-log plot fit our results to a linear model. The result is a 
  an excellent fit with r = -0.98, yielding a power law exponent of -0.79. 
  The distribution appears to be scale free with most articles having fewer than 9 
  direct First Links, while few hold most.
  \label{fig:degree distribution}
\end{figure}

The FLN's indegree exhibits a scale-free distribution where a few articles 
receive a most direct first references, while most articles receive few or none.
The average indegree for all 11 million articles is 3.6 direct First Links with a standard deviation of 89.5 links.
Only 4,826 articles have more than 100 direct First Links and $75\%$ of articles
have fewer than 9. 
When fit to a linear model on a log-log scale (log(rank) versus log(indegree)), 
the model's r value is $-0.98$ suggesting a strong log-log linear fit 
with a power law exponent of $-0.79$.

@@

\begin{figure}[tp!]
  \centering	
  \includegraphics[width=\columnwidth]{graphics/path_lengths_boxplot.png}
  \caption{
    \textbf{Path Length Distribution}
  }
  The median network depth measured by the number of First Links in a path
  is 29. The 365 cycle for "Orthodox" liturgics is the outlier to the right, while
  other historical articles about "Scotland," "UK" and so on are slightly outside
  the third quartile. More than $75\%$ of articles have path lengths between 
  $0$ and $50$ links.

@@

Of the 11 million articles, 5.5 million had an invalid link or linked back to the same article, yielding a path length of zero. 
This roughly corresponds to the official number of articles on Wikipedia: 
$~4.7$ million as of November 2014---approximately half of the 11 million 
articles in the XML dump are redirects or disambiguations, not full articles.
The most common path length is 29, with an interquartile range (26, 30).
The median path length far exceeds the popular 6-degrees of separation (see for 
example the average number of academic publications separating scholars 
\cite{six_degrees}), indicating a greater network depth. 
The network depth's depth suggests ideas are related in small independent clusters. 
Instead the many diverse ideas fit together in long-chained connections.
As a distribution, more than $75\%$ of articles have a path length below 
$50$ First Links 
while a few temporally organized paths exceed 50 links 
(see figure~\ref{fig:Path Length Distribution}).

How can we characterize the flow of ideas along a path? Next, we develop 
a metric to gauge the accumulation First Link references
and characterize a global organization 
of ideas among articles.

@@

We followed every possible path through the network, incrementing
the number of traversal visits for every First Link reference along a path 
to generate a total of 232 million traversal visits.

As a distribution, the number traversal visits per article appears to be scale-free. 
The majority of articles have fewer than 30 traversal visits. 
First Link references accumulate at a few articles.
Specifically, $99.76\%$ of articles have fewer than $100$ traversal visits; nearly $80\%$ have none. 
Meanwhile, the highest ranking 30 articles have an extremely disproportionate number of traversal visits.

\begin{figure}[tp!]
  \centering	
  \includegraphics[width=\columnwidth]{graphics/traversals_per_article.png} 
  \caption{
    \textbf{Distribution of Traversal Visits}
  }
  We fit the distribution to a linear log-log model by considering the log (base 10) transformed rank of each article against log (base 10) transformed the number of traversal visits. 
  The model explains $86\%$ of the variation in the data, yielding an r-value of $-0.93$ 
  and a power law exponent of $-0.636$. The horizontal flattening around the highest
  ranking articles is a result of the cyclic structure (see discussion on cycles).
  \label{fig:Distribution of Visits}

\end{figure}

@@

\begin{figure}[tp!]
  \centering	
  \includegraphics[width=\columnwidth]{graphics/articles_ranked.png}
  \caption{
    \textbf{highest ranking articles by number of traversal visits}
  }
  We compute the number of traversal visits for each article in the FLN (see 
  Traversal Algorithm section for details). In doing so, we can rank each article
  by the accumulation of First Links. Articles with a greater number of traversal visits
  mark greater points of First Link accumulation. The highest ranking articles by traversal visits reveals where the greatest accumulation occurs.
  \label{fig:highest visits}
\end{figure}

@@

\begin{figure}[tp!]
  \centering	
  \includegraphics[width=\columnwidth]{graphics/synonyms.png}
  \caption{
    \textbf{Number of Synonyms for the highest-ranking articles}
    We use WordNet (the largest lexical database for the English language)
    to identify the number of synonyms for an article's title. We compare 
    the distribution of synonyms of the highest-ranking 100 articles
    against all other Wikipedia articles. As the distribution demonstrates
    the highest ranking concepts have more synonyms than other articles.
  }
  \label{fig:synonyms}
\end{figure}
